---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

The following code reads the data.  Omitting records containing NA, the "aggregate" function is used to calculate the daily steps.

```{r}
activity_raw <- read.csv("activity.csv");
activity <- na.omit(activity_raw);
activity_bydate <- aggregate(activity$steps, by = list(activity$date), sum);
activity_bydate$steps <- as.numeric(activity_bydate$Group.1);
```


## What is mean total number of steps taken per day?
The histogram of the data follows:

```{r fig.width=7, fig.height=6}
hist(activity_bydate$steps, xlab= "Daily Steps", main = "Activity Histogram");
```

The mean number of daily steps is `r round(mean(activity_bydate$steps),digits=2)`.
The median value of daily steps to be `r round(median(activity_bydate$steps), digits=2)`.  This text used an inline calculation to 2 significant digits. These figures can also be found in the summary data below.


```{r}
summary(activity_bydate);
```


## What is the average daily activity pattern?

The average daily activity pattern is shown 



## Inputing missing values



## Are there differences in activity patterns between weekdays and weekends?
